Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-10-24 paper-new/concrete-relational-model.tex
edit the explanation of the perturbations for dyn

Finally, the most interesting type is of course the dynamic type. 
It may seem as though we need to define the perturbations for $D$ by
guarded recursion, but in fact we define them via least-fixpoint\footnote{
  The reason for defining the perturbations inductively is technical and has to
  do with the fact that we needed to define the perturbations corresponding to the
  predomain $\later A$ to be the perturbations of $A$, \emph{not} later of the
  perturbations of $A$. There may be a way to make this work by changing some
  definitions, but we leave this as future work}
in the category of monoids:
%
\( M_D \cong (M_{D \times D}) \oplus M_{U(D \to \li D)}. \)
%
We now explain how to interpret these perturbations as endomorphisms.
We define $i_D : M_D \to \{ f : D \to D \mid f \bisim \id \}$ by induction as follows:
%
\begin{align*}
  i_D(m_\times) &= \lambda d. \text{case $d$ of }
    \{ \itimes (d_1, d_2) \To \itimes (i_{D \times D}(m_\times)(d_1, d_2)) \alt x \To x \} \\
  i_D(m_\to) &= \lambda d. \text{case $d$ of }
    \{ \iarr (\tilde{f}) \To \iarr (\lambda t. i_{U(D \to \li D)}(m_\to)(\tilde{f}_t)) \alt x \To x \}
\end{align*}
%
In the case of a perturbation on $D \times D$, we use the interpretation for the
perturbations of $D \times D$ as defined earlier in this section, which in turn
will use $i_D$ inductively. Similarly, for a perturbation on $U(D \to \li D)$ we
use the interpretation for perturbations on $U(D \to \li D)$.
One can verify that this defines a homomorphism from $M_D \to \{ f : D \to D : f
\bisim \id \}$.

>>>>

Finally, the most interesting type is of course the dynamic type. 
It may seem as though we need to define the perturbations for $D$ by
guarded recursion, but in fact we define them via least-fixpoint
in the category of monoids:
%
\( M_D \cong (M_{D \times D}) \oplus M_{U(D \to \li D)}. \)
%
\begin{remark}
  The reason for defining the perturbations inductively is technical and has to
  do with the fact that the perturbations corresponding to the predomain $\later
  A$ need to be the same as the perturbations of $A$, \emph{not} later of the
  perturbations of $A$. There may be a way to make this work by changing some
  definitions; we leave this to future investigation.
\end{remark}
%
We now explain how to interpret these perturbations as endomorphisms.
We define $i_D : M_D \to \{ f : D \to D \mid f \bisim \id \}$ by induction as follows:
%
\begin{align*}
  i_D(m_\times) &= \lambda d. \text{case $d$ of }
    \{ \itimes (d_1, d_2) \To \itimes (i_{D \times D}(m_\times)(d_1, d_2)) \alt x \To x \} \\
  i_D(m_\to) &= \lambda d. \text{case $d$ of }
    \{ \iarr (\tilde{f}) \To \iarr (\lambda t. i_{U(D \to \li D)}(m_\to)(\tilde{f}_t)) \alt x \To x \}
\end{align*}
%
In the case of a perturbation on $D \times D$, we use the interpretation for the
perturbations of $D \times D$ as defined earlier in this section, which in turn
will use $i_D$ inductively. Similarly, for a perturbation on $U(D \to \li D)$ we
use the interpretation for perturbations on $U(D \to \li D)$.
One can verify that this defines a homomorphism from $M_D \to \{ f : D \to D : f
\bisim \id \}$.
